Structural analysis

### 2024-07-22 workflow/notebooks/draft.Rmd
added draft.rmd

# grouped_brhs <- brhs %>%
#   group_by(query) %>%
#   summarise(fs_brh = sum(fs_brh), blast_brh=sum(blast_brh), common_brh=sum(common))
blast_filtered <- filter(blast, 
                         query %in% paste0("AF-", aln_seeds, "-F1"),
                         length/qlen*100>params$coverage, 
                         length/slen*100>params$coverage,
                         evalue<as.numeric(params$eval_both)) %>% 
  group_by(query) %>% 
  slice_head(n = params$max_seqs) %>% 
  ungroup()

@@

@@

var_r2t <- scores %>% 
  filter(model!="FTPY", targets=="common") %>% 
  ggplot(aes(color=model, variance_r2t)) + 
  geom_density(size=1, alpha=.5) +
  # geom_histogram(position = "dodge") + 
  # coord_cartesian(expand = 0) +
  scale_x_continuous(limits = c(0,.2)) + 
  # facet_grid(~targets) + 
  labs(x="Variance Root-to-tip") +
  scale_color_manual(values = palettes_model) +
  theme(legend.position = 'bottom',
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,0,0,0))

>>>>

blast_filt <- filter(blast, 
                         # query %in% paste0("AF-", aln_seeds, "-F1"),
                         length/qlen*100>params$coverage, 
                         length/slen*100>params$coverage,
                         evalue<as.numeric(params$eval_both)) %>% 
  group_by(query) %>% 
  slice_head(n = params$max_seqs) %>% 
  ungroup()

fs_filt <- filter(fs, 
                      # query %in% paste0("AF-", aln_seeds, "-F1"),
                      qcov*100>params$coverage, 
                      tcov*100>params$coverage, 
                      evalue<as.numeric(params$eval_both)) %>% 
  group_by(query) %>% 
  slice_head(n = params$max_seqs) %>% 
  ungroup()

df_filt <- full_join(blast_filt, fs_filt,
                by = c("query", "target"), suffix=c("_Blast", "_Foldseek")) %>%
    filter(query!=target) %>% 
  left_join(self, by="query") %>%
  mutate(pident_Foldseek = 100*pident_Foldseek, 
         qcov_Foldseek = 100*qcov_Foldseek,
         singleton = case_when(is.na(evalue_Foldseek) ~ "only Blast",
                               is.na(evalue_Blast) ~ "only Foldseek", 
                               .default = "Common"))
# grouped_brhs <- brhs %>%
#   group_by(query) %>%
#   summarise(fs_brh = sum(fs_brh), blast_brh=sum(blast_brh), common_brh=sum(common))
blast_filtered <- filter(blast, 
                         query %in% paste0("AF-", aln_seeds, "-F1"),
                         length/qlen*100>params$coverage, 
                         length/slen*100>params$coverage,
                         evalue<as.numeric(params$eval_both)) %>% 
  group_by(query) %>% 
  slice_head(n = params$max_seqs) %>% 
  ungroup()

@@

## Fig1

```{r fig1}

fig1a <- df %>%
  group_by(singleton) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  mutate(prop=n/sum(n), Filtering="Before") %>%
  bind_rows(df_filt %>%
            group_by(singleton) %>%
            summarise(n = n()) %>%
            ungroup() %>%
            mutate(prop=n/sum(n), Filtering="After")) %>%
  mutate(Filtering=factor(Filtering, labels = c("After", "Before"), ordered = T),
         singleton=case_when(singleton=="only Blast"~"Only Bp", 
                             singleton=="only Foldseek" ~ "Only Fs",
                             .default = singleton)) %>%
  ggplot(aes(singleton, n, fill=singleton)) +
  geom_bar(aes(alpha=Filtering), stat = "identity", position = position_dodge2(reverse=TRUE)) +
  geom_text(aes(group=Filtering, label=paste0(round(prop, 2)*100, "%")),
            position = position_dodge2(reverse = TRUE, width = 0.9),
            vjust=1.3, color="white", size=3) +
  scale_fill_manual(values = palette_singleton_s, guide="none") +
  scale_alpha_manual(values = c(1,.6)) +
  scale_y_continuous(expand = c(0,0), labels = scales::label_number(scale = 1e-6)) +
  labs(fill="Class", x="Class", y="# aligned protein pairs (Mln)") +
  theme(legend.position = "bottom", legend.box="vertical")

fig1b <- lddt_df %>% 
  ggplot(aes(mean_lddt)) + 
  geom_density(fill="#BF4375", size=1) +
  scale_y_continuous(expand = expansion(mult = c(0, .1))) +
  labs(x="Mean pLDDT", y="") +
  geom_vline(xintercept = 40, linetype=2) +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.line.y = element_blank())
fig1c <- tibble(seeds) %>% 
  mutate(`Has > 4 common` = seeds %in% paste0("AF-",aln_seeds,"-F1")) %>% 
  ggplot(aes(fill=`Has > 4 common`, y=1)) + 
  geom_bar() +
  coord_cartesian(expand = 0) +
  scale_fill_manual(values = c("grey80", "black")) +
  labs(x="Has > 4 Common hits") +
  theme(legend.position = "none", 
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.line.y = element_blank())

sps <- read_delim("data/input/Hsapopi_set_sp.txt")

sptree_plot <- fortify(sptree) %>% 
  left_join(sps, by=c("label"="mnemo")) %>% 
  mutate(species=gsub("[^A-Z].+? ", ". ", species)) %>% 
  ggtree() +
  geom_tiplab(aes(label=species), nudge_x=.2, fontface="italic") + 
  geom_nodelab(color="firebrick", 
               nudge_x = -.15, hjust = 1, 
               nudge_y = .3, size=2.5)+
  scale_x_continuous(expand = expansion(mult = c(0.2, .7)))

ggsave("draft/figures/Fig1/tree.pdf", sptree_plot, width = 4, height = 4)

fig1 <- (fig1b | (fig1c/fig1d + plot_layout(heights = c(10,1)))) +
#   plot_layout(widths = c(1,2)) + 
#   plot_annotation(tag_levels = 'A') & 
#   theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))
ggsave("draft/figures/Fig1/fig1a.pdf", fig1a, width = 3, height = 4)
ggsave("draft/figures/Fig1/fig1b.pdf", fig1b, width = 4, height = 2)
ggsave("draft/figures/Fig1/fig1c.pdf", fig1c, width = 4, height = 1)

```

@@

var_r2t <- scores %>% 
  filter(model!="FTPY", targets=="common") %>% 
  ggplot(aes(fill=model, variance_r2t)) + 
  # geom_density(size=1, alpha=.5) +
  geom_histogram(position = "dodge") +
    # geom_histogram(position = "dodge") + 
  # coord_cartesian(expand = 0) +
  scale_x_continuous(limits = c(0,.2), expand = expansion(mult=0)) + 
  scale_y_continuous(expand = expansion(mult=0),
                     breaks = c(0,100,200)) +
  # facet_grid(model~.) +
  labs(x="Variance Root-to-tip") +
  scale_fill_manual(values = palettes_model) +
  theme(legend.position = 'bottom',
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,0,0,0))
